---
title: "Evaluation Platform in COPD (EPIC)"
site: "bookdown::bookdown_site"
runtime: "shiny"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
    delete_merged_file: TRUE
    table_css: FALSE
    output_dir: "docs"
    css: ["pretty-styles.css", "cayman.css", "python.css","shadow.css",  "index.html"]
    js: ["js/list-group.js", "js/code-chunk.js", "js/code-chunk-components.js"]
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: sans
        size: 2
      sharing:
        facebook: yes
        twitter: yes
        google: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'google', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    keep_tex: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

```{r, echo = FALSE, eval = TRUE}
themeColors = c("#330033", "#8cf2f2", "#c51672", "#007bff")
```
# Introduction {#introduction}

The Evaluation Platform In COPD (EPIC) was a nationally funded research project with the aim of
creating an open-source, publicly available, population-based ‘Whole Disease’ COPD model for
epidemiological projections and policy analyses in the Canadian context.

It might be helpful to take a look at the EPIC web application before reading further, as this gives a visualization of the package: [EPIC Web Application](#webapp).


