
# The Study {#thestudy}
The **epicR** package and application are based on the paper [Development and Validation of theEvaluation Platform in COPD (EPIC):A Population-Based Outcomes Modelof COPD for Canada](https://journals.sagepub.com/doi/pdf/10.1177/0272989X18824098), published in 2019.

## Background {#background}

The purpose of this study is to model how COPD (Chronic Obstructive Pulmonary Disease) affects a whole population. We combined 12 different studies to look at 6 different aspects of COPD: 

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("link-buttons.html"), collapse="\n")
HTML(rawHTML)
```

### Demographic and Risk Factor Module {-#demographic}

#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("demographic-table.html"), collapse="\n")
HTML(rawHTML)
```

### COPD Occurrence Module {-#occurence}

#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("occurrence-table.html"), collapse="\n")
HTML(rawHTML)
```

### Lung Function Module {-#lungfunction}

#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("lungfunction-table.html"), collapse="\n")
HTML(rawHTML)
```

### Exacerbation Module {-#exacerbation}

In COPD, patients can sometimes experience what are called exacerbations. 

**exacerbation**: acute worsening of COPD symptoms; acute means sudden onset, short in duration; COPD symptoms include shortness of breath, wheezing, coughing up mucus, etc.

<!-- #### Exacerbation Causes {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("exacerbation-cause.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

#### Exacerbation Severity {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("exacerbation-severity.html"), collapse="\n")
HTML(rawHTML)
```

#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("exacerbation-table.html"), collapse="\n")
HTML(rawHTML)
```

**hazard**: instantaneous exacerbation rate

### Mortality Module {-#mortality}

We categorized death into two categories: death from COPD, and death not from COPD. More formally:

**COPD-related mortality**: Death due to a severe or very severe COPD exacerbation.

**Background mortality**: Death from all other causes (excluding COPD).

#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("mortality-table.html"), collapse="\n")
HTML(rawHTML)
```

### Payoff Module {-#payoff}

Like any disease, COPD costs money to treat, and is a cost on the healthcare system. Additionally, having COPD can reduce quality of life, which depends on the severity of the diagnosis and number of exacerbations. We define cost and utility, below, and divided each into 2 categories:

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("payoff-breakdown.html"), collapse="\n")
HTML(rawHTML)
```


#### Study Summary {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("payoff-table.html"), collapse="\n")
HTML(rawHTML)
```

<!-- Chronic Obstructive Pulmonary Disease (COPD) affects 300 million people worldwide. Patients with COPD often experience what is called an **acute exacerbation**. -->





<!-- ### References {-} -->

<!-- [COPD exacerbations: definitions and classifications](https://erj.ersjournals.com/content/erj/21/41_suppl/46s.full.pdf) -->

<!-- [Toward a consensus definition for COPD exacerbations](https://www.ncbi.nlm.nih.gov/pubmed/10843984) -->

<!-- ## The Study Population {#studyPopulation} -->

<!-- The data for this study was taken from three different studies:  -->

<!-- [MACRO Clinical Trial](https://clinicaltrials.gov/ct2/show/NCT00325897) -->

<!-- [MACRO Paper](https://www.ncbi.nlm.nih.gov/pubmed/21864166) -->

<!-- [STATCOPE Clinical Trial](https://clinicaltrials.gov/ct2/show/NCT01061671) -->

<!-- [STATCOPE Paper](https://www.ncbi.nlm.nih.gov/pubmed/24836125?dopt=Abstract) -->

<!-- [OPTIMAL Clinical Trial](https://www.ncbi.nlm.nih.gov/pubmed/15611808) -->

<!-- [OPTIMAL Paper](https://annals.org/aim/fullarticle/734106) -->

<!-- ### Clinical Trial 1: MACRO {-} -->

<!-- Number of Participants: 1142 -->

<!-- Type of Study: Clinical Randomized -->

<!-- Masking: Triple (Participant, Investigator, Care-Provider) -->

<!-- #### Criteria {-} -->

<!-- - Age: > 40 years old -->
<!-- - COPD diagnosis -->
<!-- - Does not have asthma -->
<!-- - 10 pack years or more of smoking -->

<!-- #### Study Groups {-} -->

<!-- ```{r diagram, eval = TRUE, echo = FALSE} -->
<!-- library(DiagrammeR) -->
<!-- nodes <- create_node_df(n=3, type = "character", -->
<!--                         color = themeColors[1], -->
<!--                         fontcolor = "white", -->
<!--                         fillcolor = themeColors[1], -->
<!--                         shape = "oval", -->
<!--                         fontsize = 4, -->
<!--                         label = c("All \n Participants","Placebo",  -->
<!--                                   "Azithromycin \n Treatment")) -->

<!-- edges <- create_edge_df(from = c(1, 1), -->
<!--                         to = c(2, 3), -->
<!--                        rel = "leading to", -->
<!--                        color = themeColors[2]) -->

<!-- graph <- create_graph(nodes_df = nodes,  -->
<!--                       edges_df = edges)  -->
<!-- render_graph(graph, layout="tree") -->
<!-- ``` -->


<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study1-arms.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- #### Data Collection: Primary Metric {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study1-primary-metric.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- ### Clinical Trial 2: STATCOPE {-} -->

<!-- Number of Participants: 885 -->

<!-- Type of Study: Clinical Randomized -->

<!-- Masking: Triple (Participant, Investigator, Care-Provider) -->

<!-- #### Criteria {-} -->

<!-- - Age: 40 - 80 years old -->
<!-- - COPD diagnosis -->
<!-- - Does not have asthma -->
<!-- - 10 pack years or more of smoking -->

<!-- #### Study Groups {-} -->

<!-- ```{r diagram2, eval = TRUE, echo = FALSE} -->
<!-- library(DiagrammeR) -->
<!-- nodes <- create_node_df(n=3, type = "character", -->
<!--                         color = themeColors[1], -->
<!--                         fontcolor = "white", -->
<!--                         fillcolor = themeColors[1], -->
<!--                         shape = "oval", -->
<!--                         fontsize = 4, -->
<!--                         label = c("All \n Participants","Placebo",  -->
<!--                                   "Simvastatin \n Treatment")) -->

<!-- edges <- create_edge_df(from = c(1, 1), -->
<!--                         to = c(2, 3), -->
<!--                        rel = "leading to", -->
<!--                        color = themeColors[2]) -->

<!-- graph <- create_graph(nodes_df = nodes,  -->
<!--                       edges_df = edges)  -->
<!-- render_graph(graph, layout="tree") -->
<!-- ``` -->


<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study2-arms.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- #### Data Collection: Primary Metric {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study2-primary-metric.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- ### Clinical Trial 3: OPTIMAL {-} -->

<!-- Number of Participants: 432 -->

<!-- Type of Study: Clinical Randomized -->

<!-- Masking: Double (Participant, Investigator) -->

<!-- #### Criteria {-} -->

<!-- - Age: > 35 years old -->
<!-- - COPD diagnosis -->
<!-- - Does not have asthma -->
<!-- - 10 pack years or more of smoking -->
<!-- - at least 1 COPD exacerbation in the last year -->

<!-- #### Study Groups {-} -->

<!-- ```{r diagram3, eval = TRUE, echo = FALSE} -->
<!-- library(DiagrammeR) -->
<!-- nodes <- create_node_df(n=4, type = "character", -->
<!--                         color = themeColors[1], -->
<!--                         fontcolor = "white", -->
<!--                         fillcolor = themeColors[1], -->
<!--                         shape = "oval", -->
<!--                         fontsize = 4, -->
<!--                         label = c("All \n Participants","Placebo \n Inhaler",  -->
<!--                                   "Salmeteral \n Inhaler", -->
<!--                                   "Advair \n Inhaler")) -->

<!-- edges <- create_edge_df(from = c(1, 1, 1), -->
<!--                         to = c(2, 3, 4), -->
<!--                        rel = "leading to", -->
<!--                        color = themeColors[2]) -->

<!-- graph <- create_graph(nodes_df = nodes,  -->
<!--                       edges_df = edges)  -->
<!-- render_graph(graph, layout="tree") -->
<!-- ``` -->


<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study3-arms.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- #### Data Collection: Primary Metric {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("study3-primary-metric.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- ## Data Collection: Combined Trials -->

<!-- For our study, we combined the results from the three trials described above.  -->

<!-- ### Primary Metric {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("primary-metric.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- ### Secondary Metrics {-} -->

<!-- ```{r, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("secondary-metrics.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

<!-- ## Validation -->





