
# Using the Package in R {#packageinR}

## Installation

If you do not have R installed on your computer, please go to [Appendix 1](#appendix1). Once you have R installed, you can install the **epicR** package as per the instructions below:

### Windows 7 or Later {-}
1. Download and Install the latest version of R from [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/)
2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Download and Install the latest version of Rtools from [https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/) 
4. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages('devtools')
```

5. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('aminadibi/epicR')
```

### Mac OS Sierra and Later {-}
1. Download and Install the latest version of R from [https://cran.r-project.org/bin/macosx/](https://cran.r-project.org/bin/macosx/)
2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Install homebrew from [https://brew.sh](https://brew.sh) and [Xcode](https://developer.apple.com/xcode/) developer tools from the App store.
4. Open the Terminal and use brew to install `llvm`:

```{bash, eval=FALSE, echo=TRUE}
brew install llvm
```
  
5. Add the following to your `~/.bash_profile`:

```{bash, eval=FALSE, echo=TRUE}
export PATH="/usr/local/opt/llvm/bin:$PATH"
```
  
  And this to your `~/.Rprofile`:
```{bash, eval=FALSE, echo=TRUE}
Sys.setenv(PATH=paste("/usr/local/opt/llvm/bin", Sys.getenv("PATH"), sep=":"))
```
  
6. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages('devtools')
```

7. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('resplab/epicR')
```

Note: If epicR is still not compiling correctly, `gfortran` needs to be installed separately. In the terminal:

```{bash, eval=FALSE, echo=TRUE}
brew install gcc
```

Now, by default, R does not look for the HomeBrew version of gcc, so you will need to change this as well. First, you need to find the version of gcc you
are using:

```{bash, eval=FALSE, echo=TRUE}
brew list --versions gcc
```

The first number is {YOUR_GCC_MAIN_VERSION}, and the whole name is {YOUR_GCC_FULL_VERSION}. For example, on my computer, it is:

```
9.1.0
```

You will also need the folder name for your gcc, which you will need to log in as sudo to do:

```{bash, eval=FALSE, echo=TRUE}
sudo cd 
```

```{bash, eval=FALSE, echo=TRUE}
cd ~/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION}/gcc
ls
```

The folder name printed out is {YOUR_GCC_TARGET}. For example, on my computer, it is:

```
x86_64-apple-darwin18
```

In terminal, use your favourite text editor to open the file "~/.R/Makevars":

```{bash, eval=FALSE, echo=TRUE}
open ~/.R/Makevars
```

This may open a blank file, or it might have some content already. Somewhere in the file, add the following:

```
CC = gcc-{YOUR_GCC_MAIN_VERSION}
CXX = g++-{YOUR_GCC_MAIN_VERSION}
FLIBS = -L/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION}/gcc/{YOUR_GCC_TARGET}/{YOUR_GCC_FULL_VERSION} 
-L/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION} -lgfortran -lquadmath -lm
```

For example, on my computer it would be:

```
CC = gcc-9
CXX = g++-9
FLIBS = -L/usr/local/lib/gcc/9/gcc/x86_64-apple-darwin18/9.1.0
-L/usr/local/lib/gcc/9 -lgfortran -lquadmath -lm
```

Once you have done this, save the file and close the text editor. You may need to restart RStudio, and try Step 7 again.


### Ubuntu 16.04 and Later {-}
1. Install R by executing the following commands in Terminal:

```{bash, eval=FALSE, echo=TRUE}
sudo apt-key adv --keyserver keyserver.ubuntu.com --recv-keys E298A3A825C0D65DFD57CBB651716619E084DAB9
sudo add-apt-repository 'deb [arch=amd64,i386] https://cran.rstudio.com/bin/linux/ubuntu xenial/'
sudo apt-get update
sudo apt-get install r-base
```

If the installation is successful, you should be able to start R:

```{bash, eval=FALSE, echo=TRUE}
sudo -i R
```

2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Install `libcurl` from Terminal: 
```{bash, eval=FALSE, echo=TRUE}
sudo apt-get install libcurl4-openssl-dev libssl-dev
```
4. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages ('devtools')
```

5. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('resplab/epicR')
```

```{r chunk-label, eval = FALSE, echo = TRUE}
install.packages("devtools")
devtools::install_github("resplab/epicR")
```
