
# Using the Package in R {#packageinR}

## Installation

If you do not have R installed on your computer, please go to [Appendix 1](#appendix1). Once you have R installed, you can install the **epicR** package as per the instructions below:

### Windows 7 or Later {-}
1. Download and Install the latest version of R from [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/)
2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Download and Install the latest version of Rtools from [https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/) 
4. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages('devtools')
```

5. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('aminadibi/epicR')
```

### Mac OS Sierra and Later {-}
1. Download and Install the latest version of R from [https://cran.r-project.org/bin/macosx/](https://cran.r-project.org/bin/macosx/)
2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Install homebrew from [https://brew.sh](https://brew.sh) and [Xcode](https://developer.apple.com/xcode/) developer tools from the App store.

Note: `Xcode` is quite a large file, and needs about 20 GB of storage space to install. The whole process takes about an hour.

4. Open the Terminal and use brew to install `llvm`:

```{bash, eval=FALSE, echo=TRUE}
brew install llvm
```
  
5. Add the following to your `~/.bash_profile`:

```{bash, eval=FALSE, echo=TRUE}
export PATH="/usr/local/opt/llvm/bin:$PATH"
```
  
  And this to your `~/.Rprofile`:
```{bash, eval=FALSE, echo=TRUE}
Sys.setenv(PATH=paste("/usr/local/opt/llvm/bin", Sys.getenv("PATH"), sep=":"))
```
  
6. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages('devtools')
```

7. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('resplab/epicR')
```

Note: If epicR is still not compiling correctly, `gfortran` needs to be installed separately. In the terminal:

```{bash, eval=FALSE, echo=TRUE}
brew install gcc
```

Now, by default, R does not look for the HomeBrew version of gcc, so you will need to change this as well. First, you need to find the version of gcc you
are using:

```{bash, eval=FALSE, echo=TRUE}
brew list --versions gcc
```

The first number is {YOUR_GCC_MAIN_VERSION}, and the whole name is {YOUR_GCC_FULL_VERSION}. For example, on my computer, it is:

```
9.1.0
```

You will also need the folder name for your gcc, which you will need to log in as sudo to do:

```{bash, eval=FALSE, echo=TRUE}
sudo cd 
```

```{bash, eval=FALSE, echo=TRUE}
cd ~/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION}/gcc
ls
```

The folder name printed out is {YOUR_GCC_TARGET}. For example, on my computer, it is:

```
x86_64-apple-darwin18
```

In terminal, use your favourite text editor to open the file "~/.R/Makevars":

```{bash, eval=FALSE, echo=TRUE}
open ~/.R/Makevars
```

This may open a blank file, or it might have some content already. Somewhere in the file, add the following:

```
CC = gcc-{YOUR_GCC_MAIN_VERSION}
CXX = g++-{YOUR_GCC_MAIN_VERSION}
FLIBS = -L/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION}/gcc/{YOUR_GCC_TARGET}/{YOUR_GCC_FULL_VERSION} 
-L/usr/local/lib/gcc/{YOUR_GCC_MAIN_VERSION} -lgfortran -lquadmath -lm
```

For example, on my computer it would be:

```
CC = gcc-9
CXX = g++-9
FLIBS = -L/usr/local/lib/gcc/9/gcc/x86_64-apple-darwin18/9.1.0
-L/usr/local/lib/gcc/9 -lgfortran -lquadmath -lm
```

Once you have done this, save the file and close the text editor. You may need to restart RStudio, and try Step 7 again.


### Ubuntu 16.04 and Later {-}

1. Install R if you don't have it installed already. See [Appendix 1](#appendix1) for installation guide.

2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Install `libcurl` from Terminal: 
```{bash, eval=FALSE, echo=TRUE}
sudo apt-get install libcurl4-openssl-dev libssl-dev
```
4. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```{r, eval=FALSE, echo=TRUE}
install.packages ('devtools')
```

5. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github('resplab/epicR')
```

```{r chunk-label, eval = FALSE, echo = TRUE}
install.packages("devtools")
devtools::install_github("resplab/epicR")
```

## Running the Model

Now that you have installed `epicR`, you can load the library:

```{r, eval = FALSE, echo = TRUE}
library(epicR)
```

### Step 1: Initialize the Session {-}

The backend of this package is in C++. In other languages such as R/Python, memory allocation is taken care of for you. However, in C/C++, you need to allocate memory yourself. You also need to deallocate memory when the program is done. Because of this, we have the function `init_session()`.

`init_session()` is in the **core.R** file, and calls 3 C functions: `Cdeallocate_resources()`,
`Callocate_resources()`, and `Cinit_session()`. You don't really need to know how these work, but basically they allocate/deallocate memory.

You need to call this before you run anything:

```{r, eval = FALSE, echo = TRUE}
init_session()
```

In C++, a program that is successful returns `0`, so you should see this in the R console.

### Step 2: Set the Inputs {-}

If you want to set your own input parameters, I will explain how to do so in this section. If you want to use the defaults, you can skip to the next step.

The default input values are created in the file **input.R**. To set your own input values, you can use the function `createInput()` in the file **createInput.R**. Below is a list of the parameters you can set:

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("input-parameters.html"), collapse="\n")
HTML(rawHTML)
```

To create the inputs:

```{r, eval=FALSE, echo=TRUE}
input = createInput(percentFemale = 0.5,
                        timeHorizon = 20,
                        costDiscount = 0.03,
                        qalyDiscount = 0.03,
                        baselineAge = 40)
```

### Step 3: Run {-}

The `run()` function has two arguments: **max_n_agents** and **input**. If you created your own input in **Step 2**, you can put it in here; otherwise, you can leave the **input** argument blank. The **max_n_agents** argument is how many people you want in your model. The default is set to the maximum integer your computer will allow. On my computer that is `2147483647`. You can set this to a lower number if you like (it might run faster).

If you want to change the arguments:

```{r, eval=FALSE, echo=TRUE}
run(max_n_agents = 1000, input = input)
```

Otherwise:

```{r, eval=FALSE, echo=TRUE}
run()
```

### Step 4: Results {-}

Once you have run the model simulation, there are several functions to access the results. The first function is the
`get_all_events()` function. This function returns the events matrix, with every event for every patient. As you may guess,
this matrix can be quite large, depending on how many patients you ran the model with.

```{r, eval=FALSE, echo=TRUE}
eventsMatrix = get_all_events()
```

The other two functions are  `getExResults()` and `getMainResults()`. 

```{r, eval=FALSE, echo=TRUE}
extendedResults = getExResults()
mainResults = getMainResults()
```

Both functions return quite a long list of results, and as this package is in continuous development, I will refer you to the documentation in the R package for more details of how these functions work:

```{r, eval=FALSE, echo=TRUE}
??getExResults
??getMainResults
```

## Figures

We have created some figures and charts for users, although they are not exhaustive, so you may want to create your own figures as well. 

### Exacerbation Charts {-}

To look at total number of exacerbations and exacerbation rate, we have the function `plotExacerbations()`.

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("plot-exacerbations.html"), collapse="\n")
HTML(rawHTML)
```

We can look at total number of exacerbations by age group:

```{r, eval=TRUE, echo=TRUE}
plotExacerbations(nPatients = 1e4,
                  argX = "year",
                  argY = "number_of_exacerbations",
                  groupBy = "age",
                  exacType = "all", 
                  perCapita = NULL)
```

We can also look at the exacerbation rate by sex:

```{r, eval=TRUE, echo=TRUE}
plotExacerbations(nPatients = 1e4,
                  argX = "year",
                  argY = "exacerbation_rate",
                  groupBy = "sex",
                  exacType = "all", 
                  perCapita = NULL)
```

```{r, eval=TRUE, echo=TRUE}
plotExacerbations(nPatients = 1e4,
                  argX = "year",
                  argY = "exacerbation_rate",
                  groupBy = "sex",
                  exacType = "severe", 
                  perCapita = NULL)
```

### COPD Occurrence Charts {-}

To look at COPD incidence and prevalence, we have the function `plotCOPD()`.

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("plot-occurrence.html"), collapse="\n")
HTML(rawHTML)
```

We can look at COPD incidence by age:

```{r, eval=TRUE, echo=TRUE}
plotCOPD(10000, argX="age", argY="incidence", aggregateBy="none", groupBy="none")
```

We can look at COPD incidence by age group:

```{r, eval=TRUE, echo=TRUE}
plotCOPD(10000, argX="age", argY="incidence", aggregateBy="age", groupBy="none")
```

We can look at COPD incidence by year and sex:

```{r, eval=TRUE, echo=TRUE}
plotCOPD(10000, argX="year", argY="incidence", aggregateBy="none", groupBy="sex")
```

We can also look at COPD prevalence by age:

```{r, eval=TRUE, echo=TRUE}
plotCOPD(10000, argX="age", argY="prevalence", aggregateBy="none", groupBy="none")
```
